---
title: "data_handling_R_club"
author: "Robert Turnbull"
date: "17/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(catmaid)
library(dplyr)
library(tidyr)
library(elmr)
```

dplyr is the main data-handling package in R. For installation see: https://www.r-project.org/nosvn/pandoc/dplyr.html

The cheat sheet has some good information: https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

An example problem to demonstrate the uses of dplyr, focussing on generating an manipulating neuron data.


1. Read skids and neurons (in this case a selection of 10 excitatory and inhibitory mPNs)
```{r}
excitatory.mPNs.sk = catmaid_skids("Rclub_1807_Ex")
inhibitory.mPNs.sk = catmaid_skids("Rclub_1807_In")

excitatory.mPNs = read.neurons.catmaid(excitatory.mPNs.sk)
inhibitory.mPNs = read.neurons.catmaid(inhibitory.mPNs.sk)
```

2. Construct data frames with information about each neuron
```{r}
#excitatory
excitatory.df = data.frame(skid = excitatory.mPNs.sk, 
                           name = catmaid_get_neuronnames(excitatory.mPNs.sk), 
                           row.names = NULL)

#inhibitory
inhibitory.df = data.frame(skid = inhibitory.mPNs.sk, 
                           name = catmaid_get_neuronnames(inhibitory.mPNs.sk),
                           row.names = NULL)
```


3. Add some addition information about tract and type
```{r}
#the pipe operator %>% takes the output of one function and uses it as the input for another

excitatory.df %>% 
  mutate(type = "excitatory") %>% 
  mutate(tract = "mALT") -> excitatory.df

inhibitory.df %>% 
  mutate(type = "inhibitory") %>%
  mutate(tract = "mlALT") -> inhibitory.df

excitatory.df
inhibitory.df
```


These functions can be used individually, but a data frame needs specifying e.g.

mutate(df, type = "excitatory")

is equivalent to

df %>%
  mutate(type = "excitatory")
  
The latter just allows you to pipe long chains of commands together without the need to make a load of intermediary variables.


4. Merge data frames and filter by some variable
```{r}
bind_rows(excitatory.df, inhibitory.df) -> merge.df #bind_rows() joins by position

#filter by type
merge.df %>%
  filter(type == "excitatory")

#filter by tract
merge.df %>%
  filter(tract == "mlALT")

#filter by names containing "VP1"
merge.df %>%
  filter(grepl("VP1", name))
```


5. We can do more complicated transformations with relative ease in dplyr.
Here we would like to generate a connectivity matrix showing numbers of synapses between the 20 mPNs and all their downstream partners.
```{r}
connectors = catmaid_get_connector_table(skids = c(excitatory.mPNs.sk, inhibitory.mPNs.sk)) #a list of all connectors incoming and outgoing from the mPNs

connectors %>%
  filter(direction == "outgoing") %>%                #downstream targets only
  na.omit() %>%                    #removes unconnected pre-synapses
  group_by(skid) %>%                    #specify skid as a group that we want to perfrom some kind of summary on
  count(partner_skid) %>%                     #counts partner_skid per group (skid)
  spread(partner_skid, n, fill = NA) -> connectivity.matrix             #from the package "tidyr". Splits one column apart (partner_skid) and populates the new cells with new values (n)

connectivity.matrix
```
At any stage we can pipe to the View() function which creates a temporary windown with the output. Very useful when working with a long chain of commands.

