---
title: "CodeClub2"
output: html_document
date: "2024-01-22"
---
# download necessary libraries
```{r}
library(fafbseg)
library(dplyr)
```

# save data as df
```{r}
options(scipen = 999)
data <- read.delim(file='CodeClub_data.tsv')

```

# filter data by hemibrain type AVLP370 to pick a suitable neuron
```{r}
potentials <- data %>% filter(hemibrain_type=='AVLP370')
chosen1 <-potentials[1,2]
print(chosen1)


chosen1 <- "720575940613867938"
chosen1specs <- data %>%
  filter(root_id == 720575940613867938) 
```


# filter to find top 3 downstream partners
```{r}
Dstream <- flywire_partners("720575940613867938", partners = "output", details = TRUE)

Top3D <- Dstream %>%
  filter(cleft_scores > 50) %>%
  filter(post_id != pre_id)

top_3_downstream <- head(unique(Top3D$post_id), n = 3)

Dsynapse1 <- "720575940398743124"
Dsynapse2 <- "720575940633307327"
Dsynapse3 <- "720575940401353683"

Dsynapse_count <- table(Top3D$post_id)

Dsynapse1_count <- Dsynapse_count[Dsynapse1]
Dsynapse2_count <- Dsynapse_count[Dsynapse2]
Dsynapse3_count <- Dsynapse_count[Dsynapse3]

DsynapsesCount <- c(Dsynapse1_count, Dsynapse2_count, Dsynapse3_count)



# find coordinates of synapses

Dsynapse1_coords <- Top3D %>%
  filter(post_id == Dsynapse1) %>%
  select(post_x, post_y, post_z)

Dsynapse2_coords <- Top3D %>%
  filter(post_id == Dsynapse2) %>%
  select(post_x, post_y, post_z)

Dsynapse3_coords <- Top3D %>%
  filter(post_id == Dsynapse3) %>%
  select(post_x, post_y, post_z)


Dsynapses_coords <- bind_rows(
  Dsynapse1_coords %>% mutate(Synapse_ID = Dsynapse1),
  Dsynapse2_coords %>% mutate(Synapse_ID = Dsynapse2),
  Dsynapse3_coords %>% mutate(Synapse_ID = Dsynapse3)
)
                     
```
                     
# filter to find top 3 upstream partners
```{r}

Upstream <- flywire_partners("720575940613867938", partners = "input", details = TRUE)

Top3U <- Upstream %>%
  filter(cleft_scores > 50) %>%
  filter(pre_id != post_id)

top_3_upstream <- head(unique(Top3U$pre_id), n = 3)

Usynapse1 <- "720575940628567934"
Usynapse2 <- "720575940623165257"
Usynapse3 <- "720575940638077262"

Usynapse_count <- table(Top3U$pre_id)

Usynapse1_count <- Usynapse_count[Usynapse1]
Usynapse2_count <- Usynapse_count[Usynapse2]
Usynapse3_count <- Usynapse_count[Usynapse3]

UsynapsesCount <- c(Usynapse1_count, Usynapse2_count, Usynapse3_count)



# find coordinates of synapses

Usynapse1_coords <- Top3U %>%
  filter(pre_id == Usynapse1) %>%
  select(pre_x, pre_y, pre_z)

Usynapse2_coords <- Top3U %>%
  filter(pre_id == Usynapse2) %>%
  select(pre_x, pre_y, pre_z)

Usynapse3_coords <- Top3U %>%
  filter(pre_id == Usynapse3) %>%
  select(pre_x, pre_y, pre_z)


Usynapses_coords <- bind_rows(
  Usynapse1_coords %>% mutate(Synapse_ID = Usynapse1),
  Usynapse2_coords %>% mutate(Synapse_ID = Usynapse2),
  Usynapse3_coords %>% mutate(Synapse_ID = Usynapse3)
)
                     
```


# plot 3D (need to add mesh)

```{r}
#get the l2 skeleton for the main neuron
fafbseg::read_l2skel(chosen1) -> chosen1specs
```



```{r}
open3d()
#plot the main neuron and the synapses
plot3d(chosen1specs, col = 'darkred')
points3d(Dsynapses_coords, col = 'red')
points3d(Usynapses_coords, add = T, col = 'lightcoral')
plot3d(elmr::FAFB)
nview3d("frontal") #to specify the view
#nview3d("posterior")
```

*******


##create colour gradient
```{r}
# Create a vector of values
x <- c(chosen1specs, Dsynapses_coords, Usynapses_coords)

# Define a color gradient from light red to dark red
red_gradient <- colorRampPalette(c("lightcoral", "darkred"))

# Generate colors based on the gradient
mycolors <- red_gradient(3) #divide the colour gradient into 3
```

```{r}
#plot the main neuron and the synapses
plot3d(chosen1specs, col = mycolors[[3]]) #choose the darkest colour in the gradient
points3d(Dsynapses_coords, col = mycolors[[2]]) #choose the middle colour in the gradent
points3d(Usynapses_coords, add = T, col = mycolors[[1]]) #choose the lightest colour in the gradient
plot3d(elmr::FAFB)
#nview3d("frontal") #to specify the view
#nview3d("posterior")
```
# 

#opened a neuroglancer link with a flywire tab and duplicated this tab twice
#copied the ids into the appropiate tab and set the colour manually in the Render tab
https://neuroglancer-demo.appspot.com//#!gs://flyem-user-links/short/2024-01-23.081911.json

# filter data by hemibrain type AVLP390 to pick a suitable neuron
```{r}
data %>% 
  filter(hemibrain_type == "AVLP390") -> b


```

#  filter to find top 3 downstream
```{r}
flywire_partners(b, partners = 'both') -> partner_types

partner_types %>% 
  filter(cleft_scores > 50) -> partner_types

#turn the ids into characters in the dataframe
partner_types %>% 
  mutate(pre_id = as.character(pre_id)) %>% 
  mutate(post_id = as.character(post_id)) ->partner_types 
```



******************ROUGH
AVLP390neurons <- AVLP390[1:4,1]
AVLP390neurons <- c("76282330364297152", "82967841963483136", "82967704524481744", "76563942780113136")

```{r}
install.packages("viridis")

```

```{r}
# Assuming you've loaded the necessary packages

# Load the viridis package
library(viridis)

# Open 3D device
open3d()

# Plot the main neuron with a color gradient (viridis)
plot3d(chosen1specs, col = viridis(nrow(chosen1specs)))

# Add Dsynapses_coords with a red color gradient
points3d(Dsynapses_coords, col = viridis(nrow(Dsynapses_coords)))

# Add Usynapses_coords with a light coral color gradient
points3d(Usynapses_coords, col = viridis(nrow(Usynapses_coords)), add = TRUE)

# Optionally, plot additional data (elmr::FAFB)
# plot3d(elmr::FAFB)

# Set the view
nview3d("frontal")  # You can try other views like "posterior", "right", "left", etc.
```



```{r}
library(plotly)

# Chosen1 specs
mesh3d(chosen1specs)
plot_chosen1 <- plot_ly(chosen1specs, x = ~pos_x, y = ~pos_y, z = ~pos_z, type = "scatter3d", mode = "markers",
                        marker = list(color = "darkred", size = 4), name = "Chosen1 Specs")

# Dsynapses_coords
plot_Dsynapses <- plot_ly(Dsynapses_coords, x = ~post_x, y = ~post_y, z = ~post_z, type = "scatter3d", mode = "markers",
                          marker = list(color = "red", size = 4), name = "Dsynapses Coords")

# Usynapses_coords
plot_Usynapses <- plot_ly(Usynapses_coords, x = ~pre_x, y = ~pre_y, z = ~pre_z, type = "scatter3d", mode = "markers",
                          marker = list(color = "lightcoral", size = 4), name = "Usynapses Coords")

# Combine plots
combined_plot <- subplot(plot_chosen1, plot_Dsynapses, plot_Usynapses, nrows = 3)

# Show the plot
combined_plot


```
