---
title: "ALPN-Connectivity-clustering-coding-club"
author: "Anastasia Bouchorikou"
date: "2026-01-30"
output: html_document
---

#Load libraries needed

```{r}

library(fafbseg)
library(natverse)
library(dplyr)
library(tidyverse)
library(readr)
library(coconatfly)
library(coconat)
library(googlesheets4)
library(dendextend)


source("/Users/anastasiabouchorikou/aedes/2025aedes/R/funs/aedes-dataset-funs.R")
choose_aedes()

```

#Select root ids of ALPNs

```{r}

aedes=flytable_query("select * from aedes_main WHERE status NOT IN ('duplicate','fragment','tiny')")

alpns <- aedes %>%
  filter(class=="ALPN")

```

#Run connectivity clustering with upstream partners for ALPNs

```{r}

coclus_up <- cf_cosine_plot(cf_ids('/class:ALPN.*', datasets='aedes'), partners = 'in', group=TRUE, labRow = "{id}_{type}_{group}_{side}", threshold = 2, drop_dataset_prefix = T, heatmap = F)

```

##Way #1 to do it - if you dont know that there are amazing functions out there to help you 
#Extract the dendrogram from the connectivity clustering 
```{r}

ids_dend <- coclus_up$Rowv

ids_hc <- as.hclust(ids_dend)
```

#Let's decide how many clusters we want to cut the dendrogram by plotting it and then cut it into the number of choice (decide on height or cluster number)
```{r}

plot(ids_hc)

clusters <- cutree(ids_hc, h = 1.5)

order_ids <- ids_hc$labels[ids_hc$order]

```

#Fetch metadata for the clustered ids

```{r}

metadata <- cf_meta(ids = coclus_up) 

```

#On the clusters df the clustered ids are in the format ab:root_id so we need to strip the ab prefix in order to then merge the information from the clustering to the metadata df
```{r}

names(clusters) <- sub("^ab:", "", names(clusters))

add_cluster_info(metadata, coclus_up, h=1.5)

metadata <- metadata %>%
  mutate(cluster = clusters[as.character(id)])

```

#Way #2 to do it - Easier and more efficient (If you use this you should not include heatmap = TRUE in your cosine clustering. Heatmap=TRUE does not include the neuron IDs)
```{r}

metadata <- cf_meta(coclus_up)

metadata_cluster <- add_cluster_info(metadata, coclus_up, h=1.5)

```


#Assign labels
```{r}

metadata_labels <- metadata %>% 
  arrange(match(id, labels(coclus_up))) %>%
  mutate(label = paste(id, type, group, side, sep = "_"))

labels(coclus_up) <- metadata_labels$label 

```

#Plot the dendrogram with the labels
```{r}

pdf("/Users/anastasiabouchorikou/aedes/2025aedes/Anastasia/ALPN_connectivity_clustering.pdf",
    width = 40, 
    height = 10)

par(mar = c(5, 3, 1, 0), family = "Courier")

plot(coclus_up2, 
     hang = -.01,
     cex = 0.5,
     lwd = 0.6,
     main = "ALPN Connectivity Clustering")

dev.off()

```


#Generate Google spreadsheet with clustering information
```{r}

#first authenticate your google account
gs4_auth()

#then create the google sheet using the metadata df with the cluster information
ALPN_connectivity_clustering_gsheet <- gs4_create(
  name = "ALPN conectivity clustering aedes - Februrary 2025",
  sheets = list(metadata)
)

```

