---
title: "Aedes ALPN Connectivity Clustering"
author: "E Marin"
date: "2026-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(fafbseg)
library(natverse)
library(dplyr)
library(tidyverse)
library(readr)
library(coconatfly)
library(coconat)
library(googlesheets4)
library(dendextend)

source("/Users/elizabethmarin/Documents/GitHub/2025aedes/R/funs/aedes-dataset-funs.R")
```

## Retrieve Aedes metadata

```{r}
bodyann_aedes = flytable_list_rows(table = "aedes_main", base = "aedes", chunksize = 10000)
```

# Subset Aedes metadata to valid neurons

```{r}
bodyann_aedes %>%
  filter(!grepl("duplicate", status)) %>%
  filter(!grepl("fragment", status)) %>%
  filter(!grepl("tiny", status)) -> bodyann_aedes
```

# Subset to desired neurons for clustering (ALPNs)

```{r}
bodyann_aedes %>%
  filter(class %in% "ALPN") -> alpn

nrow(alpn)

alpn$root_id -> alpn_id
```

# Alternative (Marina)

```{r}
bodyann_aedes %>% 
  filter(class=="ALPN") -> alpn

alpn %>% 
  pull(root_id)->alpn_ids
```

#Are there any duplicates? (Marina)

```{r}
alpn %>% 
  filter(duplicated(alpn$root_id))
```

## Retrieve connectivity data and cluster to produce dendrogram with desired labels
# Note: Clusters look much nicer when including inputs only
# To restrict partners to class=OSN - need separate multi_connection_table

```{r}
choose_aedes()
cf_cosine_plot(
  ids = cf_ids(aedes=alpn_id),
  threshold = 2,
  partners = "inputs",
  labRow = "{id}_{type}_{group}_{side}",
  group = "type",
  heatmap = TRUE,
  matrix = FALSE,
  interactive = TRUE,
  drop_dataset_prefix = FALSE,
  keep.all.meta = TRUE,
  min_datasets = Inf,
  nas = c("zero", "drop"),
  method = c("ward.D", "single", "complete", "average", "mcquitty", "median", "centroid",
    "ward.D2")
) -> alpn_conn

# to see the interactive plot
alpn_conn

```
# Anastasia's method
```{r}
choose_aedes()
cf_cosine_plot(cf_ids("/class:ALPN", datasets='aedes'), partners = 'in', group=TRUE, labRow = "{id}_{type}_{group}_{side}") -> alpn_conn
```

# Marina's method
```{r}
cf_cosine_plot(cf_ids(aedes='class:ALPN.*'), group = 'type', partners = 'i',threshold = 2, labRow = "{id}_{group}_{type}_{side}", drop_dataset_prefix = T, heatmap = F, interactive = F) -> alpns.hc

#or use the ids instead
#cf_cosine_plot(cf_ids(aedes=alpn_ids), group = 'type', partners = 'i',threshold = 2, labRow = "{id}_{group}_{type}_{side}",drop_dataset_prefix = T, heatmap = F,interactive = F)->alpns.hc1
```

#Fetch metadata for the neurons that have been clustered - Marina

```{r}
alpn_meta = cf_meta(alpns.hc)
alpn_meta
```
# Add all the relevant clustering info (dendid, group) - Marina

```{r}
alpn_meta1 = add_cluster_info(alpn_meta, alpns.hc, h = 1.5)
alpn_meta1
```
#Rename the columns and select only the necessary ones

```{r}
alpn_meta1 %>%
  rename(root_id=id) %>% 
  rename(superclass=class) %>% 
  rename(class=subclass) %>% 
  rename(subclass=subsubclass) %>% 
  select(root_id, superclass, class, subclass)->alpn_meta2
alpn_meta2
```

#Create a Googlesheet for review

```{r}
gs4_auth()

alpns_conn <- gs4_create("Aedes ALPN conn cluster 2026-02-12")
sheet_write(alpn_meta2,ss=alpns_conn)

```
## Plot the dendrogram, changing labels to desired annotations
Plot the dendrogram.
Include root_id, type, group & side on the dendrogram labels.
Save high-resolution figures

Can use cf_meta on the hc dendrogram directly

# Can you just plot the dendrogram directly? Something like this?

```{r}
pdf(file =  ("~/aedes_alpn_mALT_nblast_dendrogram_1000_14_12_2025.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))

alpn.hc.plot<-cf_cosine_plot(
  ids = cf_ids(aedes=alpn_id),
  threshold = 2,
  partners = "inputs",
  labRow = "{id}_{type}_{group}_{side}",
  group = "type",
  heatmap = TRUE,
  matrix = FALSE,
  interactive = TRUE,
  drop_dataset_prefix = FALSE,
  keep.all.meta = TRUE,
  min_datasets = Inf,
  nas = c("zero", "drop"),
  method = c("ward.D", "single", "complete", "average", "mcquitty", "median", "centroid",
    "ward.D2"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)

dev.off()
```


# Extract the dendrogram from the connectivity clustering (Anastasia)

```{r}
ids_dend <- alpn_conn$Rowv

ids_hc <- as.hclust(ids_dend)
```

# Plot the dendrogram 
# Cut by desired number of clusters or height (Anastasia). 
# Order the ids according to the dendrogram.

```{r}
plot(ids_hc)

clusters <- cutree(ids_hc, h = 2.0)
# I opted to use a cut height, h, instead of a number of clusters, k

order_ids <- ids_hc$labels[ids_hc$order]
```
# Add annotations to the reordered list of ids to create a dataframe (Anastasia)

```{r}
metadata <- cf_meta(ids = order_ids) 
```

# Strip the ab prefix from the clusters ids so that the two dataframes can be joined (Anastasia)
# This also creates a "cluster" column in the metadata datafrome.

```{r}
names(clusters) <- sub("^ab:", "", names(clusters))

metadata <- metadata %>%
  mutate(cluster = clusters[as.character(id)])
```

# Assign desired labels from the metadata dataframe, {id}_{type}_{group}_{side} (Anastasia)

```{r}
metadata_labels <- metadata %>% 
  arrange(match(id, labels(ids_dend))) %>%
  mutate(label = paste(id, type, group, side, sep = "_"))

labels(ids_hc) <- metadata_labels$label 
```

# Save high-res figure (Anastasia)
```{r}
pdf(file =  ("~/aedes_alpn_conn_dendrogram_11_02_2026.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))

alpn.hc.plot<-plot(ids_hc, 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)

dev.off()
```

# Save high-res figure (template code)
```{r}
pdf(file =  ("~/aedes_alpn_mALT_nblast_dendrogram_1000_14_12_2025.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))
alpn.hc.plot<-plot(alpn.hc, labels=paste(aedes_alpns_ordered$root_id, aedes_alpns_ordered$side, aedes_alpns_ordered$group, aedes_alpns_ordered$axon_tract, aedes_alpns_ordered$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```

# Create a Google sheet from the dataframe (Anastasia)

https://github.com/tidyverse/googlesheets4
gs4_create() creates a brand new Google Sheet and can optionally send some initial data
sheet_write() (over)writes a whole data frame into a (work)sheet within a (spread)Sheet.

If you provide a data frame, it is used to populate the cells of a sheet and to set sheet dimensions (number of rows and columns). The header row also gets special treatment. The sheet inherits the name of the data frame, where possible.

```{r}
# Authenticate your google account
gs4_auth()

# Create the google sheet using the metadata df with the cluster information
alpns_conn <- gs4_create("Aedes ALPN connectivity clustering 2026-02-12")
sheet_write(alpn_meta2,ss=alpns_conn)
```

## Restrict the clustering to sensory neuron upstream connectivity (Marina)

```{r}
#Fetch partner connectivity data

multi_connection_table(cf_ids(aedes='class:ALPN.*'),  group = 'type', partners = 'i')->con_data
con_data
```

#Filter for sensory neuron upstream connectivity

```{r}
con_data %>% 
  filter(class=="sensory") %>% 
  cf_cosine_plot(group = 'type',threshold = 1, labRow = "{id}_{group}_{type}_{side}",drop_dataset_prefix = T, heatmap = F,interactive = F)->alpns.hc1
```
