---
title: "Aedes ALPN Connectivity Clustering"
author: "E Marin"
date: "2026-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r}
library(fafbseg)
library(natverse)
library(dplyr)
library(tidyverse)
library(readr)
library(coconatfly)
library(coconat)
library(googlesheets4)
library(dendextend)

source("/Users/elizabethmarin/Documents/GitHub/2025aedes/R/funs/aedes-dataset-funs.R")
```

## Retrieve Aedes metadata

```{r}
bodyann_aedes = flytable_list_rows(table = "aedes_main", base = "aedes", chunksize = 10000)
```

# Subset Aedes metadata to valid neurons

```{r}
bodyann_aedes %>%
  filter(!grepl("duplicate", status)) %>%
  filter(!grepl("fragment", status)) %>%
  filter(!grepl("tiny", status)) -> bodyann_aedes
```

# Subset to desired neurons (ALPNs)

```{r}
bodyann_aedes %>%
  filter(class %in% "ALPN") -> alpn

nrow(alpn)

alpn$root_id -> alpn_id
```

## Retrieve connectivity data and cluster to produce dendrogram with desired labels

```{r}
choose_aedes()
cf_cosine_plot(
  ids = cf_ids(aedes=alpn_id),
  threshold = 2,
  partners = "inputs",
  labRow = "{id}_{type}_{group}_{side}",
  group = "type",
  heatmap = TRUE,
  matrix = FALSE,
  interactive = FALSE,
  drop_dataset_prefix = FALSE,
  keep.all.meta = TRUE,
  min_datasets = Inf,
  nas = c("zero", "drop"),
  method = c("ward.D", "single", "complete", "average", "mcquitty", "median", "centroid",
    "ward.D2")
) -> alpn_conn

# Note: Clusters look much nicer when including inputs only
```
# Anastasia's method
```{r}
choose_aedes()
cf_cosine_plot(cf_ids("/class:ALPN", datasets='aedes'), partners = 'in', group=TRUE, labRow = "{id}_{type}_{group}_{side}") -> alpn_conn
```

# Marina's method
```{r}
choose_aedes()
cf_cosine_plot(cf_ids(aedes='class:ALPN.*'))
```

## Plot the dendrogram, changing labels to desired annotations
Plot the dendrogram.
Include root_id, type, group & side on the dendrogram labels.
Save high-resolution figures

# Extract the dendrogram from the connectivity clustering (Anastasia)

```{r}
ids_dend <- alpn_conn$Rowv

ids_hc <- as.hclust(ids_dend)
```

# Plot the dendrogram 
# Cut by desired number of clusters or height (Anastasia). 
# Order the ids according to the dendrogram.

```{r}
plot(ids_hc)

clusters <- cutree(ids_hc, h = 2.0)
# I opted to use a cut height, h, instead of a number of clusters, k

order_ids <- ids_hc$labels[ids_hc$order]
```
# Add annotations to the reordered list of ids to create a dataframe (Anastasia)

```{r}
metadata <- cf_meta(ids = order_ids) 
```

# Strip the ab prefix from the clusters ids so that the two dataframes can be joined (Anastasia)
# This also creates a "cluster" column in the metadata datafrome.

```{r}
names(clusters) <- sub("^ab:", "", names(clusters))

metadata <- metadata %>%
  mutate(cluster = clusters[as.character(id)])
```

# Assign desired labels from the metadata dataframe, {id}_{type}_{group}_{side} (Anastasia)

```{r}
metadata_labels <- metadata %>% 
  arrange(match(id, labels(ids_dend))) %>%
  mutate(label = paste(id, type, group, side, sep = "_"))

labels(ids_hc) <- metadata_labels$label 
```

# Save high-res figure (Anastasia)
```{r}
pdf(file =  ("~/aedes_alpn_conn_dendrogram_11_02_2026.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))

alpn.hc.plot<-plot(ids_hc, 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)

dev.off()
```

# Save high-res figure (template code)
```{r}
pdf(file =  ("~/aedes_alpn_mALT_nblast_dendrogram_1000_14_12_2025.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))
alpn.hc.plot<-plot(alpn.hc, labels=paste(aedes_alpns_ordered$root_id, aedes_alpns_ordered$side, aedes_alpns_ordered$group, aedes_alpns_ordered$axon_tract, aedes_alpns_ordered$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```

# Create a Google sheet from the dataframe (Anastasia)

https://github.com/tidyverse/googlesheets4
gs4_create() creates a brand new Google Sheet and can optionally send some initial data
sheet_write() (over)writes a whole data frame into a (work)sheet within a (spread)Sheet.

If you provide a data frame, it is used to populate the cells of a sheet and to set sheet dimensions (number of rows and columns). The header row also gets special treatment. The sheet inherits the name of the data frame, where possible.

```{r}
# Authenticate your google account
gs4_auth()

# Create the google sheet using the metadata df with the cluster information
alpn_connectivity_clusters_gsheet <- gs4_create(
  name = "Aedes ALPN connectivity clustering 2026-02-12",
  sheets = list(metadata)
)
```

