---
title: "alpn_connectivity_clustering_MG"
author: "Marina Gkantia"
date: "2026-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the libraries

```{r}
library(dplyr)
library(fafbseg)
library(coconatfly)
library(coconat)
library(googlesheets4)
library(clipr)
source("/aedes-dataset-funs.R")
choose_aedes()
```

#Retriece aedes metadata

```{r}
aedes=flytable_query("select * from aedes_main WHERE status NOT IN ('duplicate','fragment','tiny')")
```

#Filter for the ALPNs & fetch alpn rootids

```{r}
aedes %>% 
  filter(class=="ALPN")->alpns

alpns %>% 
  pull(root_id)->alpn_ids
  
```

#Are there any duplicates?

```{r}
alpns %>% 
  filter(duplicated(alpns$root_id))
```

#Run the cosine clustering
```{r}

cf_cosine_plot(cf_ids(aedes='class:ALPN.*'), group = 'type', partners = 'i',threshold = 2, labRow = "{id}_{group}_{type}_{side}",drop_dataset_prefix = T, heatmap = F,interactive = F)->alpns.hc

#or use the ids instead
#cf_cosine_plot(cf_ids(aedes=alpn_ids), group = 'type', partners = 'i',threshold = 2, labRow = "{id}_{group}_{type}_{side}",drop_dataset_prefix = T, heatmap = F,interactive = F)->alpns.hc1

```

#Fetch metadata for the neurons that have been clustered

```{r}
alpn_meta = cf_meta(alpns.hc)
alpn_meta
```

#Add all the relevant clustering info (dendid, group)

```{r}
alpn_meta1 = coconat::add_cluster_info(alpn_meta, alpns.hc, h = 1.5)
alpn_meta1
```

#Rename the columns and select only the necessary ones

```{r}
alpn_meta1 %>%
  rename(root_id=id) %>% 
  rename(superclass=class) %>% 
  rename(class=subclass) %>% 
  rename(subclass=subsubclass) %>% 
  select(-flywire_type,-instance)->alpn_meta2
alpn_meta2
```

#Create a Googlesheet for review

```{r}
gs4_auth()

alpns_conn <- gs4_create("ALPNs_connectivity_clustering_120226")
sheet_write(alpn_meta2,ss=alpns_conn)

```

#Plot the dendrogram and add labels


```{r}

pdf(file =  ("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/pn_connectivity_dendrograms/pnhc120226.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
    par(mar = c(5,3,1,0))
a<-plot(alpns.hc,labels=paste(alpn_meta2$root_id, alpn_meta2$type, alpn_meta2$group, alpn_meta2$side, sep = "_"),
         hang = -3, 
         cex = 0.5,
         lwd = 0.6)
abline(h=1.5) #add a reference line to the dendrogram showing the cut height
dev.off()
```

#Restrict the clustering to sensory neuron upstream connectivity 

```{r}
#Fetch partner connectivity data

multi_connection_table(cf_ids(aedes='class:ALPN.*'),  group = 'type', partners = 'i')->con_data
con_data
```

#Filter for sensory neuron upstream connectivity

```{r}
con_data %>% 
  filter(class=="sensory") %>% 
  cf_cosine_plot(group = 'type',threshold = 1, labRow = "{id}_{group}_{type}_{side}",drop_dataset_prefix = T, heatmap = F,interactive = F)->alpns.hc1
```

#Fetch metadata for the neurons that have been clustered

```{r}
cf_meta(alpns.hc1)->alpn_data
alpn_data
```


#Add all the relevant clustering info (dendid, group)

```{r}
alpn_data1 = coconat::add_cluster_info(alpn_data, alpns.hc1, h = 1.5)
alpn_data1
```

#Rename the columns and select only the necessary ones

```{r}
alpn_data1 %>%
  rename(root_id=id) %>% 
  rename(superclass=class) %>% 
  rename(class=subclass) %>% 
  rename(subclass=subsubclass) %>% 
  select(-flywire_type,-instance)->alpn_data2
```

#Plot the dendrogram and add labels

```{r}
pdf(file =  ("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/pn_connectivity_dendrograms/pnhc120226_osn_input.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')


    
    par(mar = c(5,3,1,0))
a<-plot(alpns.hc1, labels=paste(alpn_data2$root_id, alpn_data2$type, alpn_data2$group, alpn_data2$side, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
        lwd = 0.6)
abline(h=1.5)
dev.off()

```

#Create a Google sheet to review the clusters 

```{r}
spreadsheet<-"https://docs.google.com/spreadsheets/d/1gzkJYA0urBu1ECBPs0wOx5KzId55X6tjkYoSMg6nRu8/edit?usp=sharing"
sheet_write(alpn_data2,ss=spreadsheet, sheet = "sensory")
```

#Find out which ids are not included when you restrict the clustering to sensory neuron connectivity
```{r}
alpn_meta2 %>% 
  pull(root_id)->alpns1

alpn_data2 %>% 
  pull(root_id)->alpns2


setdiff(alpns1,alpns2) ->diff

dput(diff)

```

