---
title: "ALPN_NBLAST"
author: "Marina Gkantia"
date: "2025-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the libraries

```{r}
library(dplyr)
library(fafbseg)
library(nat.nblast)
library(nat.templatebrains)
library(dendroextras)
source("../funs/aedes-dataset-funs.R")
choose_aedes()
```


# Retrieve Aedes metadata

```{r}
aedes=flytable_query("select * from aedes_main WHERE status NOT IN ('duplicate','fragment','tiny')")
```

#Filter for ALPNs
```{r}
aedes %>% 
  filter(class=="ALPN") ->aedes_alpns
```

# are there any duplicates?
```{r}
aedes_alpns %>% filter(duplicated(aedes_alpns$root_id))
```

#Read dotprops
```{r}
alpns_dps<-read_l2dp(aedes_alpns)
```

#Check the units for dps
```{r}
boundingbox(alpns_dps)
```

# add metadata 
```{r}
alpns_dps[,]=aedes_alpns
```

#Plot ALPNs by side/Any with a wrong side annotation?
```{r}
nclear3d()
  plot3d(alpns_dps, col=case_when(side=="R" ~'red' , 
                               side=="L" ~ "green"))
```


##Mirroring registration
#using Greg's point annotations

```{r}
um="https://spelunker.cave-explorer.org/#!middleauth+https://global.daf-apis.com/nglstate/api/v1/4693107724517376"

aedes_mirrored.um=aedes_mirroreg()
```

```{r}
alpns.mirrored=xform(alpns_dps, reg = aedes_mirrored.um, subset = alpns_dps[,'side']=='L')
nclear3d()
plot3d(alpns.mirrored, col=case_when(side=="R" ~'red' , 
                               side=="L" ~ "green"))
```


#Run the NBLAST
```{r}
aba=nblast_allbyall(alpns.mirrored, .progress='text')
head(aba)
```

#Run hierarchical clustering
```{r}
aba.hc=nhclust(aba)
```

# I. Plot the neurons in the dendrogram with glue_data()
```{r}
pdf(file =  ("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/pn_nblast_dendrograms/pnhcm-121125_method1.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
    par(mar = c(5,3,1,0))
a<-plot(aba.hc, labels=glue::glue_data(aedes_alpns,"{root_id}_{side}_{group}_{type}"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)

dev.off()

```
# II. Plot the neurons in the dendrogram with glue()
```{r}
pdf(file =  ("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/pn_nblast_dendrograms/pnhcm-121125_method1.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
    par(mar = c(5,3,1,0))
a<-plot(aba.hc, labels=glue::glue("{root_id}_{side}_{group}_{type}",
                                  .envir = aedes_alpns), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)

dev.off() 

```

# III. Plot the neurons in the dendrogram with paste()
```{r}
pdf(file =  ("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/pn_nblast_dendrograms/pnhcm-121125_method1.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
    par(mar = c(5,3,1,0))

a<-plot(aba.hc, labels=paste(aedes_alpns$root_id, aedes_alpns$side, aedes_alpns$group, aedes_alpns$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```


#using Greg's point annotations + Valentina's glomeruli centroid annotations
#Extract annotations from a neuroglancer scene

```{r}
um="https://spelunker.cave-explorer.org/#!middleauth+https://global.daf-apis.com/nglstate/api/v1/4921787988049920"

mann=ngl_annotations(um)
mann
```
#Let's check that the points are all on the correct side
```{r}
nclear3d()
spheres3d(flywire_raw2nm(mann$pointA), col='magenta',radius = 2000)
spheres3d(flywire_raw2nm(mann$pointB), col='green',radius = 2000)
```
#Adjust the units and prepare the trasform
```{r}
ptsA=flywire_raw2nm(mann$pointA)
ptsB=flywire_raw2nm(mann$pointB)
aedes_mirror=nat::tpsreg(rbind(ptsA, ptsB), rbind(ptsB, ptsA))
aedes_mirror.um=nat::tpsreg(rbind(ptsA, ptsB)/1e3, rbind(ptsB, ptsA)/1e3)
```

```{r}
alpns.mirror=xform(alpns_dps, reg = aedes_mirror.um, subset = alpns_dps[,'side']=='L')
nclear3d()
plot3d(alpns.mirror, col=case_when(side=="R" ~'red' , 
                               side=="L" ~ "green"))
```

#Run the NBLAST
```{r}
aba1=nblast_allbyall(alpns.mirror, .progress='text')
head(aba1)
```

#Run hierarchical clustering
```{r}
aba.hc1=nhclust(aba1)
```

```{r}
pdf(file =  ("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/pn_nblast_dendrograms/pnhcm-121125_method2.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
    par(mar = c(5,3,1,0))

a<-plot(aba.hc1,labels=paste(aedes_alpns$root_id, aedes_alpns$side, aedes_alpns$group, aedes_alpns$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```
##Alternatively, 
# Read points transform going from Aedes to Aedes symmetric
```{r}
aedesSym <- readRDS("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/Landmarks_for_mirroring/aedes_aedesSym_1000_tps.rds")

#or you could try the trasnform with 5000 points but runs slower

aedesSym <- readRDS("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/Landmarks_for_mirroring/aedes_aedesSym_5000_tps.rds")
```

# Template provide the origin for the synapse cloud
```{r}

aedesTemplate <- nat.templatebrains::templatebrain("aedes",
                                                   voxdims = c(0.512,0.512,0.512),
                                                   BoundingBox = structure(c(35.296, 698.848, 58.784, 522.144, 0, 336.896), 
                                                                           dim = 2:3, class = "boundingbox"),
                                                   units = c("microns","microns","microns"))
```

#Filter for RHS and LHS & M ALPNs

```{r}
aedes_alpns %>% 
  filter(side=="L") ->alpnsl
aedes_alpns %>% 
  filter(side=="R") ->alpnsr
aedes_alpns %>% 
  filter(side=="M") ->alpnsm
```

#Read l2 skeletons or dotprops & add metadata

```{r}
#aedesPNs.um <- read_l2skel(aedes_alpns$root_id)/1000
 #or 
#alpnl_dps<-read_l2dp(aedes_alpns)
#or
alpnsl_dps<-read_l2dp(alpnsl) 
alpnsr_dps<-read_l2dp(alpnsr)
alpnsm_dps<-read_l2dp(alpnsm)


alpnsl_dps[,]=alpnsl
alpnsr_dps[,]=alpnsr
alpnsm_dps[,]=alpnsm

```



# Transform to Aedes symmetric
```{r}

#aedesPNsSym <-xform(alpns_dps, reg = aedesSym)

#aedesPNsSym[alpns_l]->aedesPNsSym_l
#aedesPNsSym[alpns_r]->aedesPNsSym_r
#aedesPNsSym[alpns_m]->aedesPNsSym_m
#or
alpnl_sym <- nat::xform(alpnsl_dps, reg = aedesSym)
alpnr_sym <- nat::xform(alpnsr_dps, reg = aedesSym)
alpnm_sym <- nat::xform(alpnsm_dps, reg = aedesSym)



```

# Flip LHS neurons in symmetric space
```{r}

alpnl_sym.flp <- mirror_brain(x = alpnl_sym, 
                               brain = aedesTemplate, 
                               mirrorAxis = "X", 
                               transform = "flip")

```

#Plot them 
```{r}
nclear3d()
plot3d(alpnr_sym,col="red")
plot3d(alpnl_sym.flp, col="green")
```

#Merge 3 neuronlists 
```{r}
c(alpnl_sym.flp,alpnr_sym,alpnm_sym)->all_alpns
```
#Run the NBLAST
```{r}
aba2=nblast_allbyall(all_alpns, .progress='text')
head(aba2)
```

#Run hierarchical clustering
```{r}
aba.hc2=nhclust(aba2)
```

#Reoeder the aedes_alpns to match the order of the final neuronlist

```{r}
rownames(aedes_alpns) <- aedes_alpns$root_id
aedes_alpns_ordered <- aedes_alpns[names(all_alpns), ]
```

#Plot the neurons in the dendrogram 
```{r}
pdf(file =  ("/Users/marinagkantia/Documents/GitHub/2025aedes/R/Marina/pn_nblast_dendrograms/pnhcm-121125_method3.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
    par(mar = c(5,3,1,0))

a<-plot(aba.hc2, labels=glue::glue_data(aedes_alpns_ordered,"{root_id}_{side}_{group}_{type}"), 
         hang = -.01, 
         cex = 0.5,
        lwd = 0.6)

dev.off()
```









