---
title: "NBLAST Aedes neurons"
author: "E Marin"
date: "2025-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(natverse)
library(fafbseg)
library(dplyr)
library(nat.nblast)
library(dendroextras)
#source("..R/funs/aedes-dataset-funs.R")
```

# Retrieve Aedes metadata and filter for ALPNs of appropriate status

```{r}
bodyann_aedes = flytable_list_rows(table = "aedes_main", base = "aedes", chunksize = 10000)

bodyann_aedes %>%
  filter(class %in% "ALPN") %>%
  filter(!grepl("duplicate", status)) %>%
  filter(!grepl("fragment", status)) %>%
  filter(!grepl("tiny", status)) %>%
  #filter(grepl("UNI", subclass)) %>%
  #filter(grepl("mALT", axon_tract)) %>%
  filter(!is.na(root_id)) -> aedes_alpn
nrow(aedes_alpn)

# Marina's more elegant solution:
# aedes=flytable_query("select * from aedes_main WHERE status NOT IN ('duplicate','fragment','tiny')")
# aedes %>% 
  #filter(class=="ALPN") ->aedes_alpns

table(aedes_alpn$side, useNA = 'i')
```
# Filter by side

```{r}
aedes_alpn %>%
  filter(grepl("R", side)) -> alpnr

aedes_alpn %>%
  filter(grepl("L", side)) -> alpnl

aedes_alpn %>%
  filter(grepl("M", side)) -> alpnm

table(alpnr$side, useNA = 'i')
table(alpnl$side, useNA = 'i')
```

# Retrieve neurons as simplified dotprops

```{r}
choose_aedes()
alpnr.dps = read_l2dp(alpnr)
alpnl.dps = read_l2dp(alpnl)
alpnm.dps = read_l2dp(alpnm)
```

## Perform mirroring registration including Valentina's glomerular centroids

```{r}
aedes_mirroreg <- function(units=c("microns", 'nm')) {
  um='https://spelunker.cave-explorer.org/#!middleauth+https://global.daf-apis.com/nglstate/api/v1/4921787988049920'
  mann=fafbseg::ngl_annotations(um)
  ptsA=with_aedes(fafbseg::flywire_raw2nm(mann$pointA))
  ptsB=with_aedes(fafbseg::flywire_raw2nm(mann$pointB))
  units=match.arg(units)
  if(units=='microns')
    aedes_mirror.um=nat::tpsreg(rbind(ptsA, ptsB)/1e3, rbind(ptsB, ptsA)/1e3)
  else
    aedes_mirror=nat::tpsreg(rbind(ptsA, ptsB), rbind(ptsB, ptsA))
}

alpnl.dps.m = xform(alpnl.dps, reg = aedes_mirroreg())

# Plot neurons to see whether this mirroring worked
plot3d(alpnr.dps, col = "red", lwd = 2)
plot3d(alpnl.dps.m, col = "blue", lwd = 2, add = TRUE)
```

# Combine these two neuron lists into one for NBLAST

```{r}
alpn.dps.all = c(alpnr.dps, alpnl.dps.m, alpnm.dps)
length(alpn.dps.all)
```

# Add columns for annotations that we want to feature in the dendrogram
# These two lists need to be the same length

```{r}
# Marina's solution
alpn.dps.all[,]=aedes_alpn
```

# Perform NBLAST all-by-all

```{r}
alpn.aba <- nblast_allbyall(alpn.dps.all, .progress='text')
alpn.hc <- nhclust(scoremat = alpn.aba)
```

# Reorder the aedes_alpn to match the order of the final neuronlist

```{r}
rownames(aedes_alpn) <- aedes_alpn$root_id
aedes_alpns_ordered <- aedes_alpn[names(alpn.dps.all), ]
```

# Plot the dendrogram, changing labels to desired annotations

```{r}
pdf(file =  ("~/aedes_alpn_nblast_dendrogram_14_12_2025_glom.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))
alpn.hc.plot<-plot(alpn.hc, labels=paste(aedes_alpns_ordered$root_id, aedes_alpns_ordered$side, aedes_alpns_ordered$group, aedes_alpns_ordered$axon_tract, aedes_alpns_ordered$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```


## Alternatively, use Seba's transform from point synapse cloud (5000 is much slower than 1000)

```{r}
# Read points transform going from Aedes to Aedes symmetric
aedesSym <- readRDS("/Users/elizabethmarin/Downloads/aedes_aedesSym_1000_tps.rds")
#aedesSym <- readRDS("/Users/elizabethmarin/Downloads/aedes_aedesSym_5000_tps.rds")

# Template provide the origin for the synapse cloud
aedesTemplate <- nat.templatebrains::templatebrain("aedes",
                                                   voxdims = c(0.512,0.512,0.512),
                                                   BoundingBox = structure(c(35.296, 698.848, 58.784, 522.144, 0, 336.896), 
                                                                           dim = 2:3, class = "boundingbox"),
                                                   units = c("microns","microns","microns"))
```

# Transform neurons into symmetric space and flip one side

```{r}
# Transform to Aedes symmetric
alpnr.dps.sym <- nat::xform(alpnr.dps, reg = aedesSym)
alpnl.dps.sym <- nat::xform(alpnl.dps, reg = aedesSym)
alpnm.dps.sym <- nat::xform(alpnm.dps, reg = aedesSym)

# Flip neurons of one side in symmetric space
alpnl.dps.sym.m <- nat.templatebrains::mirror_brain(x = alpnl.dps.sym, 
                                                    brain = aedesTemplate, 
                                                    mirrorAxis = "X", 
                                                    transform = "flip")

# Plot neurons to see whether this mirroring worked
clear3d()
plot3d(alpnr.dps.sym, col = "red", lwd = 2)
plot3d(alpnl.dps.sym.m, col = "blue", lwd = 2, add = TRUE)
```

# Combine the three neuron lists into one for NBLAST

```{r}
alpn.dps.all = c(alpnm.dps.sym, alpnr.dps.sym, alpnl.dps.sym.m)
length(alpn.dps.all)
```

# Add columns for annotations that we want to feature in the dendrogram

```{r}
# Marina's solution
alpn.dps.all[,]=aedes_alpn
```

# Perform NBLAST all-by-all

```{r}
alpn.aba <- nblast_allbyall(alpn.dps.all, .progress='text')
alpn.hc <- nhclust(scoremat = alpn.aba)
```

# Reorder the aedes_alpn to match the order of the final neuronlist

```{r}
rownames(aedes_alpn) <- aedes_alpn$root_id
aedes_alpns_ordered <- aedes_alpn[names(alpn.dps.all), ]
```

# Plot the dendrogram, changing labels to desired annotations

```{r}
pdf(file =  ("~/aedes_alpn_mALT_nblast_dendrogram_1000_14_12_2025.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))
alpn.hc.plot<-plot(alpn.hc, labels=paste(aedes_alpns_ordered$root_id, aedes_alpns_ordered$side, aedes_alpns_ordered$group, aedes_alpns_ordered$axon_tract, aedes_alpns_ordered$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```



# Add dendrogram columns for labels
```{r}
# Marina's solution
alpn.hc.meta=coconat::add_cluster_info(df = aedes_alpn, alpn.hc, k=40)
```



# Plot the dendrogram with labels for root_id, side, group, type

```{r}
pdf(file =  ("~/aedes_alpn_mALT_nblast_dendrogram_1000_14_12_2025.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')

par(mar = c(5,3,1,0))
plot(alpn.hc,
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```


# Marina's first solution, which did not work for me
```{r}
plot(alpn.hc, labels=glue::glue("{root_id}_{side}_{group}_{type}",
                                  .envir = alpn.hc.meta), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)

abline(h=0.8, lty='dashed')
```

# Another possible solution

```{r}
pdf(file =  ("~/aedes_alpn_mALT_nblast_dendrogram_1000.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))
alpn.hc.plot<-plot(alpn.hc, labels=paste(aedes_alpn$root_id, aedes_alpn$side, aedes_alpn$group, aedes_alpn$axon_tract, aedes_alpn$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```

## NBLAST clustering

# Alternatively try Seba's transform using skeletons

```{r}
# Read points transform going from Aedes to Aedes symmetric
aedesSym <- readRDS("/Users/elizabethmarin/Downloads/aedes_aedesSym_1000_tps.rds")

# Template provide the origin for the synapse cloud
aedesTemplate <- nat.templatebrains::templatebrain("aedes",
                                                   voxdims = c(0.512,0.512,0.512),
                                                   BoundingBox = structure(c(35.296, 698.848, 58.784, 522.144, 0, 336.896), 
                                                                           dim = 2:3, class = "boundingbox"),
                                                   units = c("microns","microns","microns"))

# Download PN skeletons and convert them to um
choose_aedes()
alpnr.um <- fafbseg::read_l2skel(alpnr$root_id)/1000
alpnl.um <- fafbseg::read_l2skel(alpnl$root_id)/1000

# Transform to Aedes symmetric
alpnr.um.sym <- nat::xform(alpnr.um, reg = aedesSym)
alpnl.um.sym <- nat::xform(alpnl.um, reg = aedesSym)

# Flip neurons of one side in symmetric space
alpnl.um.sym.flp <- nat.templatebrains::mirror_brain(x = alpnl.um.sym, 
                                                    brain = aedesTemplate, 
                                                    mirrorAxis = "X", 
                                                    transform = "flip")

# Plot neurons to see whether this mirroring worked
clear3d()
plot3d(alpnr.um.sym, col = "red", lwd = 2)
plot3d(alpnl.um.sym.flp, col = "blue", lwd = 2, add = TRUE)

# If needed transform back to original space
#aedesPNs.flp <- nat::xform(aedesPNsSym.flp, reg = reglist(aedesSym, swap=TRUE))
```

# Still need to convert these to dotprops

# Combine these two neuron lists into one for NBLAST

# Add columns for annotations that we want to feature in the dendrogram

# Perform NBLAST all-by-all

```{r}
alpn.aba <- nblast_allbyall(alpn.dps.all, .progress='text')
alpn.hc <- nhclust(scoremat = alpn.aba)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
