---
title: "NBLAST Aedes neurons"
author: "E Marin"
date: "2025-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(natverse)
library(fafbseg)
library(dplyr)
library(nat.nblast)
library(dendroextras)
#source("R/funs/aedes-dataset-funs.R")
```

# Retrieve Aedes metadata and filter for ALPNs of appropriate status

```{r}
bodyann_aedes = flytable_list_rows(table = "aedes_main", base = "aedes", chunksize = 10000)

bodyann_aedes %>%
  filter(class %in% "ALPN") %>%
  filter(!grepl("duplicate", status)) %>%
  filter(!grepl("fragment", status)) %>%
  filter(!grepl("tiny", status)) %>%
  filter(grepl("UNI", subclass)) %>%
  filter(!is.na(root_id)) -> aedes_alpn
nrow(aedes_alpn)

# Marina's more elegant solution:
# aedes=flytable_query("select * from aedes_main WHERE status NOT IN ('duplicate','fragment','tiny')")
# aedes %>% 
  #filter(class=="ALPN") ->aedes_alpns

table(aedes_alpn$side, useNA = 'i')
```
# Filter by side

```{r}
aedes_alpn %>%
  filter(grepl("R", side)) -> alpnr

aedes_alpn %>%
  filter(grepl("L", side)) -> alpnl

aedes_alpn %>%
  filter(grepl("M", side)) -> alpnm

table(alpnr$side, useNA = 'i')
table(alpnl$side, useNA = 'i')
```

# Retrieve neurons as simplified dotprops

```{r}
choose_aedes()
alpnr.dps = read_l2dp(alpnr)
alpnl.dps = read_l2dp(alpnl)
#alpnm.dps = read_l2dp(alpnm)
```

## NBLAST clustering

# Mirror one side

```{r}
aedes_mirroreg <- function(units=c("microns", 'nm')) {
  um='https://spelunker.cave-explorer.org/#!middleauth+https://global.daf-apis.com/nglstate/api/v1/4693107724517376'
  mann=fafbseg::ngl_annotations(um)
  ptsA=with_aedes(fafbseg::flywire_raw2nm(mann$pointA))
  ptsB=with_aedes(fafbseg::flywire_raw2nm(mann$pointB))
  units=match.arg(units)
  if(units=='microns')
    aedes_mirror.um=nat::tpsreg(rbind(ptsA, ptsB)/1e3, rbind(ptsB, ptsA)/1e3)
  else
    aedes_mirror=nat::tpsreg(rbind(ptsA, ptsB), rbind(ptsB, ptsA))
}

alpnl.dps.m = xform(alpnl.dps, reg = aedes_mirroreg())

# Plot neurons to see whether this mirroring worked
plot3d(alpnr.dps, col = "red", lwd = 2)
plot3d(alpnl.dps.m, col = "blue", lwd = 2, add = TRUE)
```

# Combine these two neuron lists into one for NBLAST

```{r}
alpn.dps.all = c(alpnr.dps, alpnl.dps.m)
length(alpn.dps.all)
```

# Add columns for annotations that we want to feature in the dendrogram

```{r}
# Marina's solution
alpn.dps.all[,]=aedes_alpn
```

# Perform NBLAST all-by-all

```{r}
alpn.aba <- nblast_allbyall(alpn.dps.all, .progress='text')
alpn.hc <- nhclust(scoremat = alpn.aba)
```

# Still need to change labels to desired annotations

# Add dendrogram columns for labels
```{r}
# Marina's solution
alpn.hc.meta=coconat::add_cluster_info(df = aedes_alpn, alpn.hc, k=40)
```

# Plot the dendrogram with labels for root_id, side, group, type

# Marina's first solution, which did not work for me
```{r}
plot(alpn.hc, labels=glue::glue("{root_id}_{side}_{group}_{type}",
                                  .envir = alpn.hc.meta), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)

abline(h=0.8, lty='dashed')
```

# Another possible solution

```{r}
pdf(file =  ("~/aedes_alpn_nblast_dendrogram.pdf"), 
        width = 80, 
        height = 10, 
        family = 'Courier')
    
par(mar = c(5,3,1,0))
alpn.hc.plot<-plot(alpn.hc, labels=paste(aedes_alpn$root_id, aedes_alpn$side, aedes_alpn$group, aedes_alpn$type, sep = "_"), 
         hang = -.01, 
         cex = 0.5,
         lwd = 0.6)
dev.off()
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
